Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-04-11 manuscript/manuscript.Rmd
gets spasing right around equations
adds spp biology and quadrat and record counts to methods:

\textsc{\normalsize{Andrew T. Tredennick\footnote{Corresponding author: atredenn@gmail.com}\textsuperscript{1}, Mevin B. Hooten\textsuperscript{2,3,4}, and Peter B. Adler\textsuperscript{1}}}

@@

Materials and Methods
---------------------
### Study site and data
Our demographic data come from the Fort Keogh Livestock and Range Research Laboratory in eastern Montana's northern mixed prairie near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is freely available on Ecological Archives[^link] [@Anderson2011] , and interested readers should refer to the metadata for a complete description. The site is about 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitation falling from April through September. The community is grass-dominated and we focused on the four most abundant grass species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1).

@@

We modeled each grass population based on two levels of data: individual and quadrat (Fig. 2). The individual data is the "raw" data. For the quadrat-level we data we simply sum individual basal cover for each quadrat by species. This is equivalent to a near-perfect census of quadrat percent cover because previous analysis shows that measurement error at the individual-level is small [@Chu2014]. Based on these two datasets we can compare population models built using individual-level data and aggregated, quadrat-level data.

@@

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit regressions for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately since fitting the models required different approaches. All models contain five climate covariates that we chose \emph{a priori}: "water year" precipitation at \emph{t}-1 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$ size interactions.

@@

#### Population model at the quadrat level
The statistical approach used to model aggregated data depends on the type of data collected. We have percent cover data, which can easily be transformed to proportion data. We first considered fitting three vital rate models analagous to those we fit at the individual level: one for probability of extirpation within a quadrat (analagous to survival), one for cover change within a quadrat (analagous to growth), and one for probability of colonization within a quadrat (analagous to recruitment). However, within-quadrat extirpation and colonization events were rare in our time series ($N=9$ and $N=10$, respectively, across all species). Given the broad spatial distribution of the quadrats we are studying, it is safe to assume that these events are in fact rare enough to be ignored for our purposes. So we constrained our statistical modeling of vital rates at the population level to change in percent cover within quadrats. For the remaining discussion of statistical modeling, we refer to proportion data, which is simply the percent cover divided by 100.

An obvious choice for fitting a linear model to proportion data is beta regression because the support of the beta distribution is [0,1], not including true zeros or ones. However, when we used fitted model parameters from a beta regression in a quadrat-based population model, the simulated population tended toward 100% cover for all species. We therefore chose a modeling approach based on a truncated log-normal likelihood. The model for quadrat cover change from time $t$ to $t+1$ is

@@

\begin{equation}
\text{lppd}_{\text{CV}} = \sum_{t=1}^T \text{log}_e \int [y_{t,hold}|\theta][\theta|y_{train}]d\theta,
\end{equation}

@@

\begin{equation}
\sum_{t=1}^T \text{log}_e \left(\frac{1}{S} \sum^S_{s=1}(y_{t,hold}|\theta_{ts})  \right).
\end{equation}

@@

\begin{equation}
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{w_{j}}(u_{j}))n(u_{j},t)
\end{equation}

@@

\begin{equation}
k_{j}(v_{j},u_{j},\bar{w_{j}}) = S_j(u_j, \bar{w_{j}}(u_{j}))G_j(v_{j},u_{j},\bar{w_{j}}(u_{j})) + R_j(v_{j},u_{j},\bar{w_{j}}),
\end{equation}

>>>>

\textsc{\normalsize{Andrew T. Tredennick\footnote{Corresponding author: atredenn@gmail.com}\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Mevin B. Hooten\textsuperscript{2,3,4}}}

@@

Materials and Methods
---------------------
### Study site and data
Our demographic data come from the Fort Keogh Livestock and Range Research Laboratory in eastern Montana's northern mixed prairie near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is freely available on Ecological Archives[^link] [@Anderson2011], and interested readers should refer to the metadata for a complete description. The site is about 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitation falling from April through September. The community is grass-dominated, and we focused on the four most abundant grass species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1). *B. gracilis* is a warm-season perennial grass, whereas *H. comata*, *P. smithii*, and *P. secunda* are cool-season perennial grasses. All species typically begin growth in the early spring, reach maximum growth and flower in early to mid summer (May-June), and disperse seed in mid to late summer (July-September).

@@

We modeled each grass population based on two levels of data: individual and quadrat (Fig. 2). The individual data is the "raw" data. For the quadrat-level we data we simply sum individual basal cover for each quadrat by species. This is equivalent to a near-perfect census of quadrat percent cover because previous analysis shows that measurement error at the individual-level is small [@Chu2014]. Based on these two datasets we can compare population models built using individual-level data and aggregated, quadrat-level data. At the individual level we explicitly model three vital rates: growth (13 year-to-year transitions, 29 quadrats, 18,730 records), survival (13 years, 33 quadrats, 29,353 records), and recruitment (13 years, 33 quadrats, 304 records). At the quadrat level we model population growth as change in percent cover of quadrats with non-zero cover in year *t* and in year *t-1* (13 year-to-year transitions, 19 quadrats, 784 records). For modeling population growth at the quadrat level we ignore within-quadrat extirpation and colonization events because they are very rare in our time series ($N=9$ and $N=10$, respectively, across all species). Given the broad spatial distribution of the quadrats we are studying, it is safe to assume that these events are in fact rare enough to be ignored for our purposes.

@@

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit regressions for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately since fitting the models required different approaches. All models contain five climate covariates that we chose \emph{a priori}: "water year" precipitation at \emph{t}-2 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t} (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t} (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}-1 to *t* is the transition of interest. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1).

@@

#### Population model at the quadrat level
The statistical approach used to model aggregated data depends on the type of data collected. We have percent cover data, which can easily be transformed to proportion data. An obvious choice for fitting a linear model to proportion data is beta regression because the support of the beta distribution is [0,1], not including true zeros or ones. However, when we used fitted model parameters from a beta regression in a quadrat-based population model, the simulated population tended toward 100% cover for all species. We therefore chose a modeling approach based on a truncated log-normal likelihood. The model for quadrat cover change from time $t$ to $t+1$ is

@@

\vspace{-3em}
\begin{align}
\text{lppd}_{\text{CV}} = \sum_{t=1}^T \text{log}_e \int [y_{t,hold}|\theta][\theta|y_{train}]d\theta,
\end{align}
\vspace{-3em}

@@

\vspace{-3em}
\begin{align}
\sum_{t=1}^T \text{log}_e \left(\frac{1}{S} \sum^S_{s=1}(y_{t,hold}|\theta_{ts})  \right).
\end{align}
\vspace{-3em}

@@

\vspace{-3em}
\begin{align}
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{w_{j}}(u_{j}))n(u_{j},t)
\end{align}
\vspace{-3em}

@@

\vspace{-3em}
\begin{align}
k_{j}(v_{j},u_{j},\bar{w_{j}}) = S_j(u_j, \bar{w_{j}}(u_{j}))G_j(v_{j},u_{j},\bar{w_{j}}(u_{j})) + R_j(v_{j},u_{j},\bar{w_{j}}),
\end{align}
\vspace{-3em}
